---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ImmuPop

<!-- badges: start -->
[![R-CMD-check](https://github.com/wjxiong5633/ImmuPop/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/wjxiong5633/ImmuPop/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

**ImmuPop** is an R package designed to estimate key population immunity metrics from individual serology data. The package provides the following estimators:

- **Geometric Mean Titer (GMT)**
- **Proportion of Non-Naïve Individuals**
- **Proportion of Population Immune**
- **Relative Reduction in Reproductive Number (R0)**

These estimators help assess population immunity and predict the impact of infectious diseases.

## Installation

You can install the development version of **ImmuPop** from GitHub:

```{r message=FALSE, warning=FALSE}
# install.packages("devtools")
devtools::install_github("wjxiong5633/ImmuPop")
```

## Example Usage

### Load Required Libraries

Before using the package, load **ImmuPop** and other required libraries:

```{r message=FALSE, warning=FALSE}
library(ImmuPop)
library(dplyr)
library(rlist)
library(MCMCpack)
```

### Load Raw Data

The package includes example data (`ImmuPop_raw_data`), which contains columns such as `time`, `epi`, `bsl`, `age`, and `raw_titer`.

```{r}
data("ImmuPop_raw_data")
knitr::kable(ImmuPop_raw_data, digits = 3, row.names = F)
```

### Data Preparation

Use the `generate_data()` function to preprocess raw data and categorize age groups:

```{r}
data_example <- generate_data(ImmuPop_raw_data, cut_age = c(0, 18, 50, 100))
knitr::kable(data_example, digits = 3, row.names = F) 
```

### Define Population Parameters

#### Age Proportions

Define age group proportions based on the `cut_age` parameter:

```{r}
age_prop <- c(0.2, 0.4, 0.4) # Adjust proportions according to your population
```

#### Contact Matrix

Define the **contact matrix**

```{r}
contact_matrix <- matrix(
  c(22, 16, 15,24, 28, 30, 18, 32, 35),
  nrow = 3, byrow = TRUE
)
```

#### Protection Effects

Define protection effects for children and adults based on **HAI titer levels**:

```{r}
# Maximum observed titer level should match the vector length
protect_c <- c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95)
protect_a <- c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95)
```

---

## Immunity Estimation

### Estimating Immunity at a Specific Time Point

To estimate immunity for a **specific time point** (e.g., `time = 200`):

```{r}
data_example_t <- data_example %>% filter(time == 200)

result <- ImmuPop_timet_est(
  df = data_example_t,
  protect_c = protect_c,
  protect_a = protect_a,
  age_prop = age_prop,
  contact_matrix = contact_matrix,
  sim_num = 100
)

knitr::kable(result, digits = 3, row.names = F)
```

### Estimating Immunity Across All Time Points

To estimate immunity across **all time points**:

```{r}
est_res_all <- ImmuPop_allt_est(
  df_long = data_example,
  protect_c = protect_c,
  protect_a = protect_a,
  age_prop = age_prop,
  contact_matrix = contact_matrix,
  sim_num = 100
)

knitr::kable(est_res_all, digits = 3, row.names = F)
```

### Estimating Baseline Population Immunity (Pre-Epidemic)

To estimate **baseline population immunity** before an epidemic:

```{r}
df_long_bsl <- data_example %>% filter(bsl == "yes")

est_res_bsl <- ImmuPop_bsl_est(
  df = df_long_bsl,
  protect_c = protect_c,
  protect_a = protect_a,
  age_prop = age_prop,
  contact_matrix = contact_matrix,
  sim_num = 100
)

knitr::kable(est_res_bsl, digits = 3, row.names = F)
```

