
---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
# install.packages("usethis")
# usethis::use_readme_rmd()
# usethis::use_github_action_check_standard()
```

# ImmuPop

<!-- badges: start -->
[![R-CMD-check](https://github.com/wjxiong5633/ImmuPop/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/wjxiong5633/ImmuPop/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

**ImmuPop** is an R package designed to estimate four key population immunity metrics from individual serology data:
- **Geometric Mean Titer (GMT)**

- **Proportion of Non-Naive Individuals**

- **Proportion of Population Immune**

- **Relative Reduction in Reproductive Number**

These estimators provide insights into population immunity, which can be crucial for understanding and predicting the impact of infectious diseases.

## Installation

You can install the development version of the **ImmuPop** package from GitHub using the following command:

```{r}
# install.packages("devtools")
devtools::install_github("wjxiong5633/ImmuPop")
```

## Example Usage

First, load the necessary libraries and the **ImmuPop** package:

```{r message=FALSE, warning=FALSE}
library(ImmuPop)
library(dplyr)
library(rlist)
library(MCMCpack)
```

### Sample Data

Here is an example of individual-level data that can be used to estimate immunity:

First, load your own dataset, contains columns time, epi, bsl, age,  

```{r}
data("ImmuPop_raw_data")
knitr::kable(ImmuPop_raw_data, caption = "Raw dataset", digits = 3)
```



```{r example}
data_example <- generate_data(ImmuPop_raw_data, cut_age = c(0,18,50, 100))
knitr::kable(data_example, caption = "Tidy dataset for estimation", digits = 3)
```



### Setting Age Proportions and Contact Matrix

Set the age group proportions and the contact matrix. These values should be consistent with the structure of your population.

```{r}
age_prop <- c(0.2, 0.4, 0.4)  # Proportions for age groups based on your cut age above 
contact_matrix <- matrix(c(22, 16, 15, 24, 28, 30, 18, 32,  35), nrow = 3, byrow = TRUE)
```

### Setting Protection Effects

Define the protection effect for different HAI titer levels for children and adults, make sure the length of vector is equal to the maximum titer levels.

```{r protection}
## max(data_example$titer_level) = 10
protect_c <- c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6,0.7,0.8,0.9,0.95)
protect_a <- c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6,0.7,0.8,0.9,0.95)
```


### Estimating Immunity for a Specific Time Point

To obtain immunity estimators for a specific time point (e.g., time = 282):

```{r}
data_example_t <- data_example %>% filter(time == 200)

df = data_example_t

result <- ImmuPop_timet_est(
  df = data_example_t,
  protect_c = protect_c,
  protect_a = protect_a,
  age_prop = age_prop,
  contact_matrix = contact_matrix,
  sim_num = 100
)


knitr::kable(result, caption = "Estimation at time 200", digits = 3)
```

### Estimating Immunity Across All Time Points

To obtain immunity estimates for all time points:

```{r}
est_res_all <- ImmuPop_allt_est(data_example, protect_c, protect_a, age_prop, contact_matrix, sim_num = 100)

knitr::kable(est_res_all, caption = "Estimation at each time point", digits = 3)
```

### Estimating Baseline Immunity (Pre-Epidemic)

To obtain estimators for the baseline period (pre-epidemic) for each epidemic:

```{r}
df_long_bsl <- data_example %>% filter(bsl == "yes")
est_res_bsl <- ImmuPop_bsl_est(df_long_bsl, protect_c, protect_a, age_prop, contact_matrix, sim_num = 100)
knitr::kable(est_res_bsl, caption = "Estimation at each time point", digits = 3)
```

This package provides powerful tools for estimating immunity, reduction in R0, and understanding the impact of vaccines or infections in different population groups.

---
